### 2019-01-30 intro.tex
ssr

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
As simple examples, the functions 
$f_1(x_1, x_2, x_3)$ = $x_1 \vee x_2 \vee x_3$ and
$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ are both symmetric.
This form of interchange symmetry partitions the set of variables into a set of
\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
it is $r$-symmetric, but not $(r-1)$-symmetric.
For example, the function $f_3(x_1, x_2) = x_1 \vee \overline{x_2}$
is not symmetric since $f(1, 0) \neq f(0, 1)$; however, 
it is trivially 2-symmetric.
For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
will be referred to as the \textbf{symmetry level} of $f$.
(Thus, the symmetry level of $f$ is the smallest integer $r$ such that
$f$ is $r$-symmetric.)

In the literature (see, e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
a 1-symmetric function $f$ is referred to simply
as a \textbf{symmetric} function.
Thus, if $f$ is a symmetric function, then for any
input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
Symmetric and $r$-symmetric functions for $r \geq 2$ have also been
studied in the literature in the context of discrete
dynamical systems (see, e.g., 
\cite{Barrett-etal-2007,Rosenkrantz-etal-2015,MR-2007}).

@@

The term \textbf{canalization} was first used by
Conrad Waddington (a British geneticist)
in 1942 \cite{Waddington-1942}
to express the ability of a species
to produce members with very similar observable features over a period
of time even when there are significant changes in the environment
during that period.
Since that time, the term ``canalization'' has been used to describe
the stability of a biological system with changes
in external conditions.
In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
to explain the stability of gene regulatory networks.
Kauffman observed that the use of canalyzing Boolean
functions in the model captured many observed properties of
gene regulatory networks.
The subclass of NCFs was
introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
to facilitate a rigorous analysis of the Boolean network model
for gene regulatory networks.
Many researchers have studied mathematical properties of NCFs and
pointed out the importance of NCFs in modeling biological
phenomena (e.g., \cite{Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).

There is extensive literature on the sensitivity of various classes
of Boolean functions
(e.g., \cite{Buhrman-etal-2002,Nisan-etal-1994,Zhang-2011}).
For a discussion on how the sensitivity of a Boolean function provides
an indication of its stability, the reader is referred to \cite{Kauffman-etal-2004,
Layne-2011,Layne-etal-2012}.
Li and Adeyeye \cite{Li-etal-2012} present lower and upper bounds on the
sensitivity of any NCF.
Li et al. \cite{Li-etal-2011,Li-etal-2013} conjectured that the average
sensitivity of any NCF is strictly less than  $4/3$.
This conjecture was proved by Klotz et al. \cite{Klotz-etal-2013} by establishing
a tight upper bound on the average sensitivity of any NCF.
Their methods rely on Fourier analysis of Boolean functions
\cite{Odonnell-2014}.
In \cite{Stearns-etal-2018}, 
we presented a characterization of NCFs with maximum average sensitivity.
This characterization allowed us to present an alternative proof 
of the upper bound in \cite{Klotz-etal-2013} and derive a closed
form expression for the number of such NCFs.

>>>>

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
As a simple example, each pair of variables in the function 
$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric.
This form of interchange symmetry partitions the set of variables into a set of
\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
it is $r$-symmetric, but not $(r-1)$-symmetric.
For example, the function $f_3(x_1, x_2) = x_1 \vee \overline{x_2}$
is not symmetric since $f(1, 0) \neq f(0, 1)$; however, 
it is trivially 2-symmetric.
For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
will be referred to as the \textbf{symmetry level} of $f$.
(Thus, the symmetry level of $f$ is the smallest integer $r$ such that
$f$ is $r$-symmetric.)

In the literature (see, e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
a 1-symmetric function $f$ is referred to simply
as a \textbf{symmetric} function.
Thus, in any symmetric function, each pair of variables is symmetric.
As a simple example, the function $f_2(x_1, x_2, x_3)$ = 
$x_1 \oplus x_2 \oplus x_3$ (defined above) is symmetric.
If $f$ is a symmetric function, then for any
input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
Symmetric and $r$-symmetric functions for $r \geq 2$ have also been
studied in the literature in the context of discrete
dynamical systems (see, e.g., 
\cite{Barrett-etal-2007,Rosenkrantz-etal-2015,MR-2007}).

@@

The term \textbf{canalization} was first used by
Conrad Waddington (a British geneticist)
in 1942 \cite{Waddington-1942}
to express the ability of a species
to produce members with very similar observable features over a period
of time even when there are significant changes in the environment
during that period.
Since that time, the term ``canalization'' has been used to describe
the stability of a biological system with changes
in external conditions.
In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
to explain the stability of gene regulatory networks.
Kauffman observed that the use of canalyzing Boolean
functions in the model captured many observed properties of
gene regulatory networks.
The subclass of NCFs was
introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
to facilitate a rigorous analysis of the Boolean network model
for gene regulatory networks.
It is known that the class of NCFs coincides with that
of unate cascade functions \cite{}
Many researchers have studied mathematical properties of NCFs and
pointed out the importance of NCFs in modeling biological
phenomena (e.g., \cite{Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
Additionally, properties of NCFs such as sensitivity and 
stability have also been studied
by many researchers 
\cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
Li-etal-2011,Li-etal-2013,Klotz-etal-2013,
Stearns-etal-2018}. 
To our knowledge, the relationship between NCFs and symmetric
Boolean functions has not been studied in the literature.
